Nathan Greenslit Data Analysis Project

### 2023-04-04 code/analysis_code/dust/daily_dusts.qmd
edit
Clean scripts

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---
**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection. This utilizes a data set that has the variations of dust trends **
Libs
```{r}
library(here)
library(tidyverse)
```

Data
```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```
**There are a few columns that we do not need or that contain NaN. So let's remove those.**

Clean
```{r}
dust<- dust %>%
  select(date,t1,t7,t13,t19,tsum,t7sum) %>%
  filter(!is.na(dust$date))
```

Plot T7sum
```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()
```

>>>>

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---

**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection (t7sum). This utilizes a data set that has the variations of dust trends.**

# Getting Started

#### Libraries

```{r}
library(here)
library(tidyverse)
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```

# Cleaning

There are a few columns that we do not need or that contain NaN. So let's remove those.

#### Remove NAs

```{r}
dust<- dust %>%
  filter(!date %in% as.Date('2022-07-04'))
```

#### Select Dates for Daily Time series

```{r}
dust<-
  dust %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19'))
```

### Plot T7sum over Daily Time series

<!--# Make this graph look nicer -->

```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()
```

### 2023-04-04 code/analysis_code/dust/dust_and_copies.qmd
Clean scripts

---
title: "Dust_Copies_Plot"
format:
  html:
    theme: default
---
**This script is good for visualizing copies per mL data and dust data.**
Library
```{r}
library(here)
library(tidyverse)
library(ggpubr)
```

Load Data 
```{r}
dust_sum<-read_csv(here("data", "processed_data", "dust_master.csv")) 
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
```

Daily Data frames
```{r}
daily_dust_sum<- dust_sum %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

@@

Plot Daily Dust Concentration
```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = t7sum),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 10))

@@

Total Daily Vibrio
```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

@@

Combine Dust and Copies
```{r}
fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

@@

Save as PNG
```{r}
png(file = here("results", "plots", "vib_dust_daily.png"))
fig
dev.off()
```

>>>>

---
title: "Dust_Copies_Plot"
format:
  html:
    theme: default
---

**This script is good for visualizing copies per mL data and dust data.**

# Getting Started

#### Libraries

```{r}
library(here)
library(tidyverse)
library(ggpubr)
```

#### Load Data

```{r}
dust_sum<-read_csv(here("data", "processed_data", "dust_master.csv")) 
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
```

# Cleaning

#### Select Daily Data frames

```{r}
daily_dust_sum<- dust_sum %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

@@

# Visualization

#### Plot Daily Dust Concentration

```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = t7sum),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 10))

@@

#### Total Daily Vibrio

```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

@@

#### Combine Dust and Copies

```{r}
fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

@@

#### Save as PNG

```{r}
png(file = here("results", "plots", "vib_dust_daily.png"))
fig
dev.off()
```

### 2023-04-04 code/wrangling_code/2_total_copies_calculator.qmd
Clean scripts

---
title: "Copies/mL"
format:
  html:
    theme: default
---
*This script is for cleaning qPCR results on Total Vibrio enumeration. It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio. The data set contains a daily and monthly time series. We will be examining daily.*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
data <-read_csv(here("data", "raw_data", "copies_master.csv"))
```

Select and Filter for Daily and Monthly Time Series 
```{r}
daily<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Daily")

@@

Create Site Column from Sample Name 
```{r}
site_daily<- daily %>%
  mutate(site = substr(daily$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)

@@

Add Dilution Column and Filter column. The dilution column applies if samples were diluted prior to qPCR. For example, if I diluted a sample 1:10, my dilution column would denote "10".  The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter prior to qPCR. 
```{r}
dilution_daily<- site_daily %>%
  mutate(dilution = case_when(
                          Sample == "BO7"~"10", #BO site was diluted 1:10
                          Sample == "BO8"~"10",
                          Sample == "BO9"~"10",
                          Sample == "BO11"~"10",
                          Sample == "BO12"~"10",
                          Sample == "BO13"~"10",
                          Sample == "BO14"~"10",
                          Sample == "BO15"~"10",
                          Sample == "BO16REDO"~"1",
                          Sample == "BO17"~"10",
                          Sample == "BO18"~"10",
                          Sample == "BO19"~"10",
                              site  == "C2"~ "1", #Canals and Gulf were not diluted
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

@@

Go from CQ --> Copies/mL
```{r}
copies_daily<- filter_daily %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ):  This comes from running a standard curve during qPCR. 
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume: Each qPCR reaction was 10µL
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume: Each 10µL reaction had 2.5µL of DNA added. 
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

@@

Some samples did not amplify: Replace NaN with 0 for copies/mL
```{r}
replace_daily<- copies_daily %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve

@@

Date Column: This adds a date correpsonding the a number in the sample name. 
```{r}
date_daily<- replace_daily %>%
  mutate(day = substr(replace_daily$Sample,3,4)) %>% #Creates new column `day` 
  mutate(date = case_when(day == "7"~"2022-07-07", #Makes date based off of day
                          day == "7R"~"2022-07-07",
                         day ==  "8"~"2022-07-08",
                         day ==  "9"~"2022-07-09",
                         day ==  "11"~"2022-07-11",
                         day ==  "12"~"2022-07-12",
                         day ==  "13"~"2022-07-13",
                         day ==  "14"~"2022-07-14",
                         day ==  "15"~"2022-07-15",
                         day ==  "16"~"2022-07-16",
                         day ==  "17"~"2022-07-17",
                         day ==  "18"~"2022-07-18",
                         day ==  "19"~"2022-07-19")) %>%
  select(copies_mL,site,date,Type)

@@

Each sample (Date) was done in triplicate for qPCR: We need to take the average of qPCR triplicates 
```{r}
final_daily<- date_daily %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

@@

Combine Daily and Monthly data sets back togthether. 
```{r}
copies_final<- full_join(final_monthly,final_daily)
summary(copies_final)
str(copies_final)
```

Save as RDS and CSV
```{r}
#saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
write.csv(copies_final,file= here("data","processed_data","total_clean_copies.csv"))

>>>>

---
title: "Copies/mL"
format:
  html:
    theme: default
---

**This script is for cleaning qPCR results on Total Vibrio enumeration. It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio. The data set contains a daily and monthly time series. We will be examining daily.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
```

#### Data

```{r}
data <-read_csv(here("data", "raw_data", "copies_master.csv"))
```

# Cleaning

#### Select and Filter for Daily and Monthly Time Series

```{r}
daily<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Daily")

@@

#### Create Site Column from Sample Name

```{r}
site_daily<- daily %>%
  mutate(site = substr(daily$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)

@@

#### Add Dilution Column and Filter column.

The dilution column applies if samples were diluted prior to qPCR. For example, if I diluted a sample 1:10, my dilution column would denote "10". The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter upon collection.

```{r}
dilution_daily<- site_daily %>%
  mutate(dilution = case_when(
                          Sample == "BO7"~"10", #BO site was diluted 1:10
                          Sample == "BO8"~"10",
                          Sample == "BO9"~"10",
                          Sample == "BO11"~"10",
                          Sample == "BO12"~"10",
                          Sample == "BO13"~"10",
                          Sample == "BO14"~"10",
                          Sample == "BO15"~"10",
                          Sample == "BO16REDO"~"1",
                          Sample == "BO17"~"10",
                          Sample == "BO18"~"10",
                          Sample == "BO19"~"10",
                              site  == "C2"~ "1", #Canals and Gulf were not diluted
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

@@

#### Go from CQ --> Copies/mL

```{r}
copies_daily<- filter_daily %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ):  This comes from running a standard curve during qPCR. 
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume: Each qPCR reaction was 10µL
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume: Each 10µL reaction had 2.5µL of DNA added. 
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

@@

#### Replace NaN with 0 for copies/mL

Some samples did not amplify and have no values:

```{r}
replace_daily<- copies_daily %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve

@@

#### Data Column

This adds a date corresponding the a number in the sample name.

```{r}
date_daily<- replace_daily %>%
  mutate(day = substr(replace_daily$Sample,3,4)) %>% #Creates new column `day` 
  mutate(date = case_when(day == "7"~"2022-07-07", #Makes date based off of day
                          day == "7R"~"2022-07-07",
                         day ==  "8"~"2022-07-08",
                         day ==  "9"~"2022-07-09",
                         day ==  "11"~"2022-07-11",
                         day ==  "12"~"2022-07-12",
                         day ==  "13"~"2022-07-13",
                         day ==  "14"~"2022-07-14",
                         day ==  "15"~"2022-07-15",
                         day ==  "16"~"2022-07-16",
                         day ==  "17"~"2022-07-17",
                         day ==  "18"~"2022-07-18",
                         day ==  "19"~"2022-07-19")) %>%
  select(copies_mL,site,date,Type)

@@

#### qPCR Triplicated --> Average

Each sample (Date) was done in triplicate for qPCR (technical replicates): We need to take the average of qPCR triplicates

```{r}
final_daily<- date_daily %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

@@

#### Combine Daily and Monthly data sets back together.

```{r}
copies_final<- full_join(final_monthly,final_daily)
summary(copies_final)
str(copies_final)
```

#### Save as CSV

```{r}
#saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
write.csv(copies_final,file= here("data","processed_data","total_clean_copies.csv"))

### 2023-04-04 products/manuscript/Manuscript.qmd
stats
Added LM for Nutrient Data
Clean scripts

@@

@@

### Cross-correlation Analysis

@@

>>>>

# Questions for Dr. Handel:

1.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?

    -   Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.

    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into,

@@

<!--# Let's add Nutrient of interest Data -->

@@

### Check Distributions:

<!--# Add distribution statement -->

### Cross-correlation Analysis:

@@

<!--# Need to update these plots -->
